---
title: "R Notebook"
output: html_notebook
---

This notebook loads and explores a dataset of texts produced by important
organizations that have been a part of the first and second wave of feminist 
movement. The idea is to explore the text using quantitative analysis and explore
the performance of several unsupervised machine learning methods in identifying
which text belongs to which wave of feminism.
This is not the data I am working on for my project; that data is restricted.
This file was created as a summary of steps taken in the actual analysis,
simplyfing the steps and changing the data, but preserving the logic and
methods used. 


#Part 1: Data cleaning

```{r message = FALSE, warning=FALSE}
library(readr)
library(magrittr)
library(stringi)
library(dplyr)
dataset <- read_csv(
  "../data/raw/comparativewomensmovement_dataset.csv",
  local = locale(encoding = "latin1")) %>%
  as.data.frame()
```

Removing redundant variables and checking the dataset structure:
```{r}
dataset <- dataset[, -c(1, 2)]

str(dataset)
```

Proceed to check if all texts are in English, using the textcat library.
First, the text is lowercased:
```{r message=FALSE, Warning=FALSE}
library(textcat)

language_eval_string <- subset(dataset, 
                               select = c(text_string)) %>%
  mutate(lowercase_text_string = tolower(text_string))

```

Textcat provides a number of texts for identification. However, limiting the 
options is advisable, since the identification with all available languages
creates false positives (e.g. identifying English as Afrikaans, etc). Testing
with English, German and French as possible options
```{r}
my_profiles <- TC_byte_profiles[names(TC_byte_profiles) %in% c("english", "german", "french")]
my_profiles

doc_lang_eval <- textcat(language_eval_string$lowercase_text_string, 
                         p = my_profiles) %>%
  as.data.frame()

doc_lang_eval <- as.data.frame(doc_lang_eval)

table(doc_lang_eval)
```

To evaluate, we bind this to the dataset and check for the 
documents that were labeled as "NA" and "German". A quick glimpse shows
that "NA" observations had an empty text field, while the german one
was in fact just a letter K. Thus, these are filtered out:
```{r}
dataset_lang_eval <- cbind(dataset,
                           doc_lang_eval)

dataset_lang_eval_filtered <-  filter(dataset_lang_eval, . == "english")
```
Save the file:
```{r}
write_csv(dataset_lang_eval_filtered, "../data/processed/dataset_lang_eval_filtered.csv")
```